### 2020-02-07 jfm/draft_p2_jfm.tex
Update abstract following Emmanuel's comments

\documentclass{jfm}
\pdfoutput=1
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage[svgnames]{xcolor}
\usepackage{graphicx}
\usepackage{acro}
\usepackage{subfig}
\usepackage{caption}

\usepackage[normalem]{ulem}
\usepackage{soul}

\graphicspath{{../figures/}}

\captionsetup[figure]{justification=raggedright}

\newcommand{\set}[1]{\ensuremath{\mathcal{#1}}}

\DeclareAcronym{pdf}{
	short=PDF,
	long=Probability Density Function,
}
\DeclareAcronym{iid}{
	short=i.i.d.,
	long=Independent Identically Distributed
}
\DeclareAcronym{ou}{
	short=OU,
	long=Ornstein-Uhlenbeck,
}
\DeclareAcronym{gktl}{
	short=GKTL,
	long=Giardina-Kurchan-Tailleur-Lecomte,
}
\DeclareAcronym{ams}{
	short=AMS,
	long=Adaptive Multilevel Splitting,
}
\DeclareAcronym{scgf}{
	short=SCGF,
	long=Scaled Cumulant Generating Function,
}
\DeclareAcronym{lbm}{
	short=LBM,
	long=Lattice Boltzmann Method,
}
\DeclareAcronym{lbe}{
	short=LBE,
	long=Lattice Boltzmann Equation,
}
\DeclareAcronym{lgca}{
	short=LGCA,
	long=Lattice Gas Cellular Automata,
}
\DeclareAcronym{lbgk}{
	short=LBGK,
	long=Lattice Bhatnagar-Gross-Krook
}
\DeclareAcronym{OU}{
	short=O-U,
	long=Ornstein--Ulhenbeck
}
\DeclareAcronym{dns}{
	short=DNS,
	long=Direct Numerical Simulation,
}
\DeclareAcronym{md}{
	short=MD,
	long=Molecular Dynamics,
}
\DeclareAcronym{cfd}{
	short=CFD,
	long=Computational Fluid Dynamics,
}

\definecolor{myred}{HTML}{b30000}
\definecolor{mygreen}{HTML}{008000}
\newcommand{\EL}[1]{{\color{myred}{#1}}}
\newcommand{\TL}[1]{{\color{mygreen}{#1}}}

\newcommand{\ZZ}[1]{{\color{magenta}{#1}}}

\title{
  Numerical study of extreme mechanical efforts exerted by a turbulent flow on a bluff body by direct and rare-event sampling techniques
}

\author{Thibault Lestang\aff{1}\aff{2}
  \corresp{\email{thibault.lestang@cs.ox.ac.uk}},
  Freddy Bouchet\aff{1}
  \and Emmanuel L\'evêque\aff{2}}
\affiliation{\aff{1}Univ Lyon, ENS de Lyon, Univ Claude Bernard de Lyon, CNRS, Laboratoire de Physique, F-69342 Lyon, France
\aff{2}Univ Lyon, Ecole Centrale de Lyon, Univ Claude Bernard de Lyon, INSA de Lyon, CNRS, Laboratoire de M\'ecanique des Fluides et d'Acoustique, F-69134 Ecully cedex, France}
\begin{document}

\maketitle

\begin{abstract}
  This study investigates, by means of numerical simulations, extreme mechanical efforts exerted by a turbulent flow impinging onto a bluff body, and examines the relevance of statistical algorithms to efficiently sample these events.
  The drag experienced by a squared obstacle placed in a turbulent channel flow (in two dimensions) is taken as a representative case study.
  Direct sampling shows that extreme fluctuations are closely related to the presence of a strong vortex blocked in the near wake of the obstacle.
  Two different algorithms are considered to speed up the sampling of such flow scenarii related to extreme values of the drag, namely the \ac{ams} and the \ac{gktl} algorithms.
  The idea behind these algorithms is to replace a long simulation by a set of much shorter ones, running in parallel, with dynamics that are replicated or pruned to sample large-amplitude events more frequently.
  These two algorithms are representative of the two main class of algorithms applicable to non-equilibrium and deterministic dynamics.
  Whilst the \ac{ams} is general, the \ac{gktl} algorithm is particularly suited to sample trajectories based on the average value of a given observable.
  These algorithms have been shown to be relevant for a wide range of problems in statistical physics, computer science, biochemistry, but it is here the first application to a fluid-structure interaction problem.
  Practical evidence is given that the fast sweeping time of fluid structures past the obstacle has a strong
  influence on the efficiency of the selection-and-cloning rules of these two algorithms. While
  the \ac{ams} algorithm does not yield significant run-time savings as compared to direct
  sampling, the \ac{gktl} algorithm appears to be efficient to sample extreme fluctuations of
  the time-averaged drag and estimate related statistics such as return times.
\end{abstract}

\include{introduction}

\include{section_description_case_study}

\include{section_extremes_direct_sampling}

\include{rare_events_algorithms}

\section{Conclusion}
\label{conlusion}

In a first part, we showed that the phenomenology describing extreme drag fluctuations is very different from typical events.
By means of a long simulation of reference, we observed that such extreme events are caused by the (temporary) trapping of vorticity very close to the base of the square.
The lifetime of extreme drag fluctuations of the order of the turnover time, and the corresponding drag signal is very peaked around these extreme values.
Furthermore, this lifetime is of the order of the correlation timescale of the drag and therefore each successive extreme drag events are decorrelated for one another.
From a timeseries spanning four million turn-over times we also illustrated that the phenomenology of extremes of the time-averaged drag cannot be reduced to a well defined scenario.
Indeed, extreme average values result both from rare successions of typical fluctuations, of the occurrence a fewer number of rare extreme fluctuations of the instantaneous drag.

In order to extend our study, and consider its replication on more complex flows, we considered the application of two popular rare-event sampling algorithms.
Because of the short lifetime of extreme drag fluctuations, the \ac{ams} does not lead to a better sampling
when using the drag itself as a score function.
The choice of the score function is difficult for complex dynamics, and optimal score function is problem-dependant.
To enable the application of the \ac{ams} to a wide range of complex dynamical systems, an important question is the systematic estimation of the optimal score function, for instance using machine learning approaches.

We obtained more satisfactory results for the averaged drag using the \ac{gktl} algorithm.
Indeed, using the algorithm we could sample trajectories corresponding to values of the time-averaged drag unreachable from a direct sampling of similar computational cost.
Qualitatively, we found the phenomenology of the sampled extreme events to be consistent with the findings of the part of the study.
However, a more quantitative analysis is required to assess the ability of the \ac{gktl} algorithm to equally sample the different dynamics leading to a given extreme value of the average drag.
Nonetheless, our study shows that the sampled ensemble is useful to compute estimators over extreme events, such as return times of extreme fluctuations of the average drag.

\section{Acknowledgements}
The authors thank Francesco Ragone, Corentin Herbert, Charles-Edouard Bréhier and Eric Simonnet for useful discussions and suggestions on various aspects of this work.
T.L and F.B acknowledge support from the European Research Council under the European Union's seventh Framework Program (FP7/2007-2013 Grant Agreement No. 616811).
Simulations have been performed on the local HPC facilities at École Normale Supérieure de Lyon (PSMN) and École Centrale de Lyon (PMCS2I).
The facilities at PSMN are supported by the Auvergne-Rhône-Alpes region (GRANT CPRT07-13 CIRA) and the national Equip@Meso grant (ANR-10-EQPX-29-01).
Finally, the authors would like to thank the three anonymous referees for their detailed review and insightfull comments and suggestions.

\appendix

\include{appendix_lbm}
\include{appendix_ams}
\include{appendix_gktl}
\include{appendix_perturbation}

\bibliographystyle{jfm}
\bibliography{../biblio}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: