### 2020-02-14 jfm/appendix_perturbation.tex
Update appendix on perturbation

\section{Perturbation at branching time}
\label{app:perturb_branching_time}
As mentioned in sections~\ref{sec:ams} and~\ref{sec:gktl}, the application of the \ac{ams} and the cloning algorithm to deterministic systems requires the addition of a perturbation to the initial condition of
resampled trajectories.
In the absence of any perturbation, the resampling would not yield new trajectories, but instead exact copies of the trajectories the resampling is based on.

The state of the flow at any point in time is given by the velocity field $\mathbf{u}$ and the pressure $p$, where the dependence on both space and time of both variables is implicit.
A general strategy for perturbing the state of the flow is to construct a perturbative divergence-free field $\delta \mathbf{u}$ and a corresponding $\delta p$ solution of the
Poisson equation.
In the context of the \ac{lbm} , the state of the system is described at a mesoscopic level by the ensemble the populations $\{f_k(x_j, y_j)\}_{0\le k \le 8, 0 \le i <N_x, 0 \le j < N_y}$ (see appendix~\ref{app:lbm}).
As a consequence, it is difficult to obtain the mesoscopic state from the knowledge of the macroscopic state $(\mathbf{u}+\delta \mathbf{u}, p + \delta p)$ from with the \ac{lbm} model can be initialised.

In this work, we work around this issue by designing the perturbation directly at the level of the mesoscopic populations.
At each lattice site $\mathbf{x}$, the populations $\{f_i\}_{0\le i \le 8}$ are replaced by
\begin{equation}
  f_{i}(\mathbf{x}) \longrightarrow f_i(\mathbf{x}) + \epsilon \sum_{n=1}^{N} \alpha_{n}f_{i}^{n}(\mathbf{x}), \,\,\, 0\leq i \leq 8,
  \label{eq:perturb_pop}
\end{equation}
with each member of the set $\{\alpha_n\}_{1\leq m \leq N}$ drawn uniformly from $[0,1]$ and $\epsilon$ the amplitude of the perturbation. The perturbed populations are then rescaled so that no mass is added in the system.
For more details concerning this procedure, see appendix C of~\cite{lestang:tel-01974316}.
In other words, the perturbation is computed as a linear combination of $N$ solutions $\{f_i^{(n)}(\mathbf{x})\}_{1\le n \le N}$ of the numerical model~\eqref{eq:lbe}.
Note that the Lattice Boltzmann Equation~\eqref{eq:lbe} is not linear, and therefore a linear combination of solutions cannot be assimilated as a solution of the numerical model.
However, nonlinearities in~\eqref{eq:lbe} are located in the equilibrium distributions~\eqref{eq:lbe_eq} and are of order of the Mach number $\mathbf{u}/c_s$ that is kept small compared to $1$, see section~\ref{sec:test_flow}.

In this work, we chose $N = 10$ and $\epsilon = 0.002$.
The value of $\epsilon$ was chosen so that the perturbation does not impact the statistics of the drag.
In order to check this assumption, we generated a timeseries of duration $T_{tot} = 10^5 \tau_c$, and introduced a perturbation with a period $\tau = \tau_c / 2$, mimicking the perturbation of clones in the cloning algorithm.
Figure~\ref{fig:pdf_drag_with_perturbation} shows that the correct statistics are obtained even in presence of this slight perturbation.

\begin{figure}
  \centering
  \includegraphics[width=.7\linewidth]{pdf_drag_with_perturbation/pdf_drag_with_perturbation}
  \caption{Estimation of the \ac{pdf} with and without periodic perturbation.
    Note that the \ac{pdf} with perturbation is based on a timeseries that is ten times shorter ($T_{tot} = 10^5 \tau_c$) than the one used to estimate the blue \ac{pdf} ($T_{tot} = 10^6 \tau_c$), hence the aspect of the tails of the former.}
  \label{fig:pdf_drag_with_perturbation}
\end{figure}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "draft_p2_jfm"
%%% End: